---
title: "Final"
output: html_document
date: "`r Sys.Date()`"
author: "Joe Montoya"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Steps
```{r install, echo=FALSE}
install.packages("BiocManager")
BiocManager::install("devtools")
BiocManager::install("affy")
BiocManager::install("limma")
## The following is the annotation used for the data 
## we are using. It will be different for other data.
## Run annotate script to ensure it is correct.
BiocManager::install("ath1121501.db")
```

# Import libraries  
- library("affy")
- library("ath1121501.db")
- library("limma")  

Note: annotation library ath1121501.db is used because it is specified

```{r libraries, include=FALSE}
library("affy")
library("ath1121501.db")
library("limma")
```

# Load data  
This example uses .cel file from [here](https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E-GEOD-10323?query=mycorrhizal%20fungi)

```{r load, include=FALSE}
eset<-ReadAffy(celfile.path="data")
slotNames(eset)
eset
class(eset)
dim(eset)
str(eset)
annotation(eset)
sampleNames(eset)
head(geneNames(eset))
head(probeNames(eset))
```

Extract phenotype information and other

```{r extract, echo=FALSE}
edata <- exprs(eset)
fdata <- fData(eset)
pdata <- pData(eset)
```

Read data

```{r read, echo=FALSE}
head(edata)
head(fdata)
head(pdata)
```

# A look at distributions of raw data

```{r boxplot1, echo=FALSE}
boxplot(eset, target = "core", names=pData(eset)$sample,
        main = "Boxplot of log2-intensities - raw data")
```

# Correct data using rma and plot again

```{r rma, include=FALSE}
eset <- rma(eset)
```

```{r boxplot 2, echo=FALSE}
boxplot(exprs(eset), target = "core", names=pData(eset)$sample,
        main = "Boxplot of normalized log2-intensities")
```

# Add feature data for ath1121501.db probe

```{r feature, include=FALSE}
anno <- select(ath1121501.db,
               keys = (featureNames(eset)),
               columns = c("SYMBOL", "GENENAME"),
               keytype = "PROBEID")
```

```{r annotate, echo=FALSE}
head(anno)
anno<-anno[!is.na(anno$SYMBOL),]
anno<-anno[table(anno$PROBEID)==1,]
```

Add expression set

```{r expression, include=FALSE}
fData(eset)<- anno[match(featureNames(eset),anno$PROBEID),]
head(fData(eset))
```

Add phenotypic information

```{r add pheno, include=FALSE}
fData(eset)<- anno[match(featureNames(eset),anno$PROBEID),]
head(fData(eset))
pData(eset)$Treatment<-factor(c(rep("DMSO",2),rep("CA",2)))
pData(eset)
```

Colapse probes and remove genes without names

```{r collapse, include=FALSE}
exprs.gene<-avereps(exprs(eset),ID=fData(eset)$SYMBOL)
exprs.gene<-exprs.gene[!is.na(rownames(exprs.gene)),]
```

Design model matrix, fit linear model

```{r design fit, include=FALSE}
design <- model.matrix(~0 +pData(eset)$Treatment)
colnames(design) <- levels(pData(eset)$Treatment)
design
fit <- lmFit(exprs.gene,design)
fit
```

Fit to contrast matrix and Bayes

```{r fit again, include=FALSE}
contrast.matrix<- makeContrasts(CA-DMSO,levels=levels(pData(eset)$Treatment))
contrast.matrix
fit <- contrasts.fit(fit, contrast.matrix)
fit <- eBayes(fit)
fit
```

# Generate volcano plot

```{r volcanoplot, echo=TRUE}
volcanoplot(fit, coef = 1, highlight=10, names=rownames(fit))
```

# Show the top 10 genes differentially expressed between treatment groups

```{r other, echo=TRUE}
topTable(fit,adjust="BH",n=10,coef=1)
DEG<-topTable(fit, coef=1, adjust.method="BH", sort.by="p", p.value=0.05)
head(DEG)
```
